---
title: "Homework #2"
author: "Name: Nicholas Ang"
output: html_document
---


## Question 1  


__(a)__ Calculate the sum of 6 and 12.

```{r error=T} 
sum(6,12)

```

__(b)__ Calculate 5 plus 5.

```{r error=T} 
5+5

```

__(c)__ Subtract 3 from 6.

```{r error=T} 
6-3
```

__(d)__ Multiply 2 times 5.

```{r error=T} 
2*5

```

__(e)__ Calculate 12 divided by 3.

```{r error=T} 
12/3

```

__(f)__ Calculate $2^5$.

```{r error=T} 
2^5

```


## Question 2


__(a)__ Calculate $|-10|$. 

```{r error=T} 
abs(-10)

```

__(b)__ Calculate $\sqrt{7}$.

```{r error=T} 
sqrt(7)

```

__(c)__ Calculate $\log(9)$.

```{r error=T} 
#ln(9) or log base e(9)
log(9)

#log base 10 (9)
log10(9)

#log base 2 (9)
log2(9)

```

__(d)__ Calculate $e^3$.

```{r error=T} 
exp(3)
```


## Question 3


__(a)__ Complete the following R code to define (3, 3, 3)

```{r error=T} 
c(  3,  3,  3)
```


__(b)__  Complete the following R code to define (3, 3, 3)

```{r error=T} 
rep( 3, 3)
```


__(c)__ Complete the following R code to define (4, 4, 4, 4)

```{r error=T} 
c(  4, 4 , 4 , 4 )
```


__(d)__ Complete the following R code to define (4, 4, 4, 4)

```{r error=T} 
rep(4 ,4 )
```


__(e)__ Complete the following R code to define (2, 4, 6)

```{r error=T} 
c( 2 , 4 , 6 )
```


__(f)__ Complete the following R code to define (2, 4, 6)

```{r error=T} 
seq(2, 6 , by = 2)
```


__(g)__ Complete the following R code to define (1, 3, 5)

```{r error=T} 
c( 1 , 3 , 5 )
```


__(h)__ Complete the following R code to define (1, 3, 5)

```{r error=T} 
seq( 1 , 5 , by = 2)
```


__(i)__ Complete the following R code to define (2, 2, 3, 3, 3, 4, 5, 5, 5)

```{r error=T}  
rep( 2:5, c(2,3,1,3))
```


__(j)__ Complete the following R code to define (1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3)

```{r error=T} 
rep( 1:3, 4)
```


__(k)__ Complete the following R code to define (6, 6, 7, 7, 5, 5)

```{r error=T} 
rep(c(6,7,5) , each = 2)
```


__(l)__ Complete the following R code to define (0, 0.01, 0.02, $\ldots$, 0.99, 1)

```{r error=T} 
seq( 0 ,  1, by = 0.01)
```


## Question 4  


__(a)__ Complete the following R code to define

\[
\mathbf{x}=\left(\begin{array}{c}
4\\
3\\
1
\end{array} \right)
\]

```{r error=T} 
x <- c(4,3,1)
print(x)
```


__(b)__ Complete the following R code to define

\[
\mathbf{y}=\left(\begin{array}{r}
-1\\
2\\
1
\end{array} \right)
\]

```{r error=T} 
y <- c(-1,2,1)
print(y)
```


__(c)__ Calculate $3\mathbf{x}$.

```{r error=T} 
3*x

```


__(d)__ Calculate $\mathbf{x}+\mathbf{y}$.

```{r error=T} 
x+y

```


__(e)__ Calculate $\mathbf{x}^\prime \mathbf{y}$.

```{r error=T} 
t(x)%*%y
```


__(f)__ Complete the following R code to define

\[
\mathbf{A} =
\left(\begin{array}{rrr}
3 & -1 & 2 \\ 1 & 5 & 4 \\ 3 & 2 & -1
\end{array} \right)
\]

```{r error=T} 
A <- matrix(c(3,1,3,-1,5,2,2,4,-1),3,3)
print(A)
```


__(g)__ Complete the following R code to define

\[
\mathbf{B} =
\left(\begin{array}{rr}
2 & 0 \\ 1 & -1 \\ 2& 1
\end{array} \right)
\]

```{r error=T} 
B <- matrix(c(2,0,1,-1,2,1),nrow = 3, ncol = 2, byrow = TRUE)
print(B)
```


__(h)__ Complete the following R code to define

\[
\mathbf{C} =
\left(\begin{array}{rr}
4 & 2 \\ 3 & 1
\end{array} \right)
\]

```{r error=T} 
C <- matrix(c(4,3,2,1),2,2)
print(C)
```


__(i)__ Use 'diag' function to define

\[
\mathbf{D} =
\left(\begin{array}{rrr}
5 & 0 & 0 \\ 0 & 4 & 0 \\ 0 & 0 & 2
\end{array} \right)
\]

```{r error=T} 
D <- diag(c(5,4,2))
print(D)
```


__(j)__ Calculate $\mathbf{AB}$.

```{r error=T} 
A%*%B

```


__(k)__ Calculate $\mathbf{BA}$. If you get an error message, explain why.

```{r error=T} 
B%*%A

```
The error in (k) is caused by the differences in rows and columns in A and B. While they can matrix multiply AB, it cannot happen with BA. A is a 3x3 matrix while B is a 3x2 matrix. Since the number of columns of A is the same as the number of rows of B, they can multiply. BA cannot be multiplied since the B is 3x2 and A is 3x3 so the number of columns in B is not the same as the number of rows in A. This results in an error since the matrices cannot be multiplied.


__(l)__ Calculate $\mathbf{A}^\prime$.

```{r error=T} 
t(A)

```


__(m)__ Calculate $(\mathbf{AB})^\prime$.

```{r error=T} 
t((A%*%B))

```


__(n)__ When we compute the inverse of a matrix, we need to check whether its determinant is zero or not. If the determinant is zero, the inverse matrix does not exist. 
Let's check whether the determinant of $\mathbf{C}$ is zero or not. To calculate the determinant, we can use `det` function. For example, the determinant of `a` can be found by using `det(a)`.

```{r error=T} 
det(C)
```
The inverse matrix of C exists since the determinant of c is not 0.

__(o)__ Calculate $\mathbf{D}^{-1}$.

```{r error=T} 
solve(D)

```


__(p)__ Calculate $\mathbf{ABC}$.

```{r error=T} 
A%*%B%*%C
```

__(q)__ Calculate $(\mathbf{A+D})^\prime$.

```{r error=T} 
t((A+D))

```

__(r)__ Calculate $(\mathbf{A-2D})^{-1}$.

```{r error=T} 
solve((A-(2*D)))

```

__(s)__ Complete the following R code to define

\[
\mathbf{E} =
\left(\begin{array}{rrrr}
1 & 0 & 0 & 0 \\ 0 & 1 & 0 & 0 \\ 0 & 0 & 1 & 0\\ 0 & 0 & 0 & 1
\end{array} \right)
\]

```{r error=T} 
E <- diag(4)
print(E)
```


## Question 5  

We define the stock prices of Apple as `apple` below.

```{r error=T} 
apple <- c(109.49, 109.90, 109.11, 109.95, 111.03, 112.12, 113.95, 113.30, 115.19, 115.19,115.82, 115.97, 116.64, 116.95, 117.06, 116.29, 116.52, 117.26, 116.76, 116.73, 115.82)
```


__(a)__ Find the length of `apple`.

```{r error=T} 
length(apple)

```


__(b)__ Find the third element of `apple`.

```{r error=T} 
apple[3]

```


__(c)__ Sort `apple` in decreasing order and save it as `apple2`.

```{r error=T} 
apple2 <- sort(apple, decreasing = TRUE)
print(apple2)

```


__(d)__ From `apple2`, find the second highest stock price.

```{r error=T} 
apple2[2]

```
Since it is sorted from highest to lowest, 2nd highest is the 2nd element
