---
title: "Homework #4"
author: "Name: Nicholas Ang"
output: html_document
---


## Question 1  


__(a)__  Create a $3\times 4$ matrix `x` using the following vector of 18 values: `c(10,5,1,4,7,4,3,3,1,3,4,3)`.

```{r error=T} 
x = matrix(c(10,5,1,4,7,4,3,3,1,3,4,3),3,4)
print(x)
```


__(b)__  Using `x`, identify the dimension-specific index positions of elements that are either 3 OR 4.

```{r error=T} 
which(x == 3 | x == 4, arr.ind = T)

```


__(c)__  Replace all elements in `x` that are less than 3 OR greater than or equal to 7 with the value 100.

```{r error=T} 
x[which(x < 3 | x >= 7)] = 100
print(x)

```


## Question 2


Suppose that you are leading a research team of five data analysts and that you want to evaluate their performance. 
To do this, you track their speed, evaluate each analyst as `"slow"`, `"medium"` or `"fast"`, and save the results in `speed_vector`.


__(a)__  Assign `speed` a vector with 5 entries, one for each analyst. Each entry should be either `"slow"`, `"medium"`, or `"fast"`. 
Analyst 1 is medium, Analyst 2 is slow, Analyst 3 is slow, Analyst 4 is medium, and Analyst 5 is fast.

```{r error=T} 
speed <- c("medium","slow","slow","medium","fast")
print(speed)
```


__(b)__  The categories of `speed` have a natural ordering. Convert `speed` to an ordered factor vector.

```{r error=T} 
speed <- factor(speed, ordered = T, levels = c("slow","medium","fast"))
print(speed)
```


__(c)__  Let us summarize the performance rating of the five analysts. Find how many analysts received the rating of `"slow"`, `"medium"`, and `"fast"`, respectively, by using an R function.

```{r error=T} 
summary(speed)

```


__(d)__  Having a bad day at work, Analyst 2 enters your office and starts complaining that Analyst 5 is slowing down the entire project. 
To check if his statement is true, you decide to compare the performance rating of Analyst 2 to that of Analyst 5. 
Is Analyst 2 faster than Analyst 5? Use a relational operator.

```{r error=T} 
speed[2] > speed[5]

```


## Question 3


In a small town, we recorded the ages of COVID-19 patients as below. 

```{r} 
age <- c(50,34,63,23,75,23,12,45,3,34,64,43,13,66,34,45,46,49,23,63,88)
```

Instead of actual ages, we will use the following classification of age groups. 

Group 1: younger than 10, Group 2: 10-19, Group 3: 20-29, Group 4: 30-39, 
Group 5: 40-49, Group 6: 50-59, Group 7: 60-69, Group 8: 70-79, Group 9: 80 or older

Using `cut` function, save the age group into `age_grouped`.

```{r error=T} 
#closed parenthesis on right, open parenthesis on left
age_grouped <- cut(age, breaks = c(0,10,20,30,40,50,60,70,80,Inf), right = FALSE)
print(age_grouped)
```


## Question 4


__(a)__  We have the following variables: `my_numeric`, `my_character`, `my_logical`.

```{r error=T} 
my_numeric <- 42
my_character <- "some text"
my_logical <- TRUE
```

Change `my_numeric` to a character variable, `my_character` to a logical variable, `my_logical` to a numeric variable.

```{r error=T} 
as.character(my_numeric)
as.logical(my_character)
as.numeric(my_logical)

```


__(b)__  If you got an error message or NA in part (a), explain why.

You get an error message because there is no conversion from some text to a logical value of TRUE or FALSE. It can only convert to logical if R knows what it means. For example, "TRUE" or "T" can be converted from character to logical but words like "some text" cannot be converted since R does not know what "some text" means.

## Question 5


In this question, we will use a data set `mtcars` that is already stored in R. 
To use such a data set, we type

```{r} 
data(mtcars)
```

This data comprises fuel consumption and some aspects of automobile design and performance for several automobiles.  
To see the first few observations, we use `head()` function:

```{r} 
head(mtcars)
```


__(a)__  To see the structure of `mtcars`, use `str()` function.

```{r error=T} 
str(mtcars)

```
 
 
__(b)__  From part (a), find the number of observations ($n$) and the number of variables ($p$). Complete the following sentence.

The number of observations is 32 and the number of variables is 11.


__(c)__  Check if `mtcars` is a vector, matrix, array, factor, or data frame. 

```{r error=T} 
is.vector(mtcars)
is.matrix(mtcars)
is.array(mtcars)
is.factor(mtcars)
is.data.frame(mtcars)

```


## Question 6


You want to construct a data set that describes the main characteristics of eight planets in our solar system. 
The main features of a planet are:

`name`: Mercury, Venus, Earth, Mars, Jupiter, Saturn, Uranus, Neptune

`type`: The type of planet (Terrestrial or Gas Giant)

`diameter`: The planet's diameter relative to the diameter of the Earth.

`rotation`: The planet's rotation across the sun relative to that of the Earth.

`rings`: If the planet has rings or not (TRUE or FALSE).

```{r} 
name <- c("Mercury", "Venus", "Earth", 
          "Mars", "Jupiter", "Saturn", 
          "Uranus", "Neptune")
type <- c("Terrestrial planet", 
          "Terrestrial planet", 
          "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", 
          "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, "ten", 0.532, 
              11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 
              0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
```


__(a)__  Construct a data frame `planets` with `name`, `type`, `diameter`, `rotation`, and `rings`, in this order.

```{r error=T} 
planets <- data.frame(name, type, diameter, rotation, rings)
planets2 = data.frame(planets, name, type, diameter, rotation, rings)
print(planets)
print(planets2)
mylist = list(planets,planets2,name,type,diameter,rotation,rings)
print(mylist)
df = data.frame(mylist)
df = data.frame(diag(3))
df
as.list(planets)
```

From now on, we will use only a data frame `planets`. To avoid any confusion, let's remove the variables `name`,`type`, `diameter`, `rotation`, `rings`.
For example, we will be able to use `planets$name`, but not `name`, after removing the globally defined variables.

```{r} 
rm(list=c("name","type", "diameter", "rotation", "rings"))
```


__(b)__  Use `str()` function to investigate the structure of `planets`.

```{r error=T} 
str(planets)

```


__(c)__  From part (b), determine if each variable is numeric, character, or logical. 

`name` is a character variable. 

`type` is a character variable. 

`diameter` is a numeric variable. 

`rotation` is a numeric variable. 

`rings` is a logical variable. 


__(d)__  Find the diameter of Mercury. (Use R code)

```{r error=T} 
planets$diameter[planets$name == "Mercury"]

```


__(e)__  Find all data on Mars (the fourth row).


```{r error=T} 
planets[planets$name == "Mars",]

```


__(f)__  Find the first five values in the `diameter`.

```{r error=T}  
head(planets$diameter, 5)

```


__(g)__  Use `$` sign to select the `rings` variable from `planets`.

```{r error=T} 
planets$rings

```


__(h)__  Some planets in our solar system have rings and others do not. Unfortunately you cannot recall their names. 
Find the names of the planets that have rings by using square brackets and part (g).

```{r error=T} 
planets$name[planets$rings]

```


__(i)__  In part (h), we were interested in only names. Now let us find all information available (that is, all five variables) for the planets with rings. Complete the following R code.

```{r error=T}  
planets[planets$rings,   ]
```


__(j)__  Using `subset` function, repeat part (i).

```{r error=T}  
subset(planets, planets$rings)
```


__(k)__  Using `subset` function, find the planets that have a diameter smaller than Earth. That is, find planets with `diameter` < 1. 

```{r error=T} 
subset(planets, planets$diameter < planets$diameter[planets$name == "Earth"])

```


__(l)__  Arrange a data frame `planets` such that it starts with the smallest planets and ends with the largest one. That is, arrange `planets` in ascending order of `diameter`. 

```{r error=T}  
planets[order(planets$diameter, decreasing = FALSE), ]

```


## Question 7


A library `MASS` has a data set `road` with the annual deaths in road accidents for half the US states. 

```{r} 
library(MASS)
head(road)  # see the first few observations
```

Rows show 'state' name.
Columns are
`deaths` number of deaths.
`drivers` number of drivers (in 10,000s).
`popden` population density in people per square mile.
`rural` length of rural roads, in 1000s of miles.
`temp` average daily maximum temperature in January.
`fuel `fuel consumption in 10,000,000 US gallons per year.
     

__(a)__  Find the number of observations ($n$) and the number of variables ($p$). 

```{r error=T} 
str(road)

```
The number of observations is 26 and the number of variables is 6.

__(d)__  Rearrange `road` in descending order of `deaths` and save it as `road_by_deaths`. Print out the first six observations of `road_by_deaths`.

```{r error=T} 
road_by_deaths = road[order(road$deaths, decreasing = TRUE),]
head(road_by_deaths)

```


__(c)__  Find a state that has the highest annual fuel consumption. 
(Hint: Arrange `road` in descending order of `fuel`) 

```{r error=T} 
road_by_fuel = road[order(road$fuel,decreasing = TRUE),]
rownames(head(road_by_fuel,1))

```


__(d)__  Find the number of drivers in the state with the lowest number of deaths. 
(Hint: Arrange `road` in ascending order of `deaths`) 

```{r error=T} 
road_by_deaths_asc = order(road$deaths)
road$drivers[road_by_deaths_asc[1]]
```


## Question 8 


Suppose we have two data frames, the first containing information on plant life forms and the second containing information of time of flowering. 
Both data frames contain variables for genus name and species name.

```{r} 
lifeforms <- data.frame(Genus=c("Acer", "Acer", "Ajuga", "Conyza", "Lamium"), 
                        species=c("platanoides", "palmatum", "reptans", "sumatrensis", "album"),
                        lifeform=c("tree","tree", "herb", "annual", "herb"))

flowering <- data.frame(Genus=c("Acer", "Ajuga", "Brassica", "Chamerion", "Conyza", "Lamium"), 
                        species=c("platanoides", "reptans", "napus", "angustifolium", "bilbaoana", "album"),
                        flowering=c("May","June", "April", "July", "August", "January"))
print(lifeforms)
print(flowering)
```

We produced a single data frame showing information on both life form and flowering time. 
We included all the species, with missing values (NA) inserted when flowering times or life forms are not known,

```{r eval=F} 
      Genus       species lifeform flowering
1      Acer      palmatum     tree      <NA>
2      Acer   platanoides     tree       May
3     Ajuga       reptans     herb      June
4  Brassica         napus     <NA>     April
5 Chamerion angustifolium     <NA>      July
6    Conyza     bilbaoana     <NA>    August
7    Conyza   sumatrensis   annual      <NA>
8    Lamium         album     herb   January
```

Use `merge` function to get the output as above.

```{r error=T} 
merge(lifeforms,flowering, all = TRUE)

```


## Question 9


__(a)__  Create a list `mylist` that contains, `var1`, `var2`, `var3`, and `var4`, in this order.

`var1`: a sequence of 20 evenly spaced numbers between -4 and 4.

`var2`: a $3\times 3$ matrix with logical values as follows:

\[
\left(\begin{array}{rrr}
F & T & T \\ T & F & T \\ T & F & F
\end{array} \right)
\]

`var3`: a character vector with the two strings `"STAT 446"` and `"MATH 646"`.

`var4`: an ordered factor vector containing the observations `c("LOW", "MED", "LOW", "MED", "MED", "HIGH")`.
The levels have a magnitude order with `"LOW"` < `"MED"` < `"HIGH"`.

```{r error=T} 
var1 <- seq(-4,4,length.out = 20)
var2 <- matrix(c(F,T,T,T,F,F,T,T,F), 3,3)
var3 <- c("STAT 446", "MATH 646")
var4 <- factor(c("LOW", "MED", "LOW", "MED", "MED", "HIGH"), ordered = T, levels = c("LOW", "MED", "HIGH"))
mylist <- list(var1 = var1,var2 = var2,var3 = var3,var4 = var4)
names(mylist) <- c("var1", "var2", "var3", "var4")  # named list
print(mylist)
rm(list=c("var1", "var2", "var3", "var4"))  # for study purpose, remove variables and use only mylist
```


__(b)__  Find all values from `var1` that are greater than 1. 

```{r error=T} 
mylist$var1[mylist$var1 > 1]

```


__(c)__  Extract the second row of the second and third columns of `var2`.

```{r error=T} 
mylist$var2[2,2:3]

```


__(d)__  Change the second element of `var3` to `"STAT 646"`.

```{r error=T} 
mylist$var3[2] = "STAT 646"
print(mylist$var3)

```


__(e)__  Using `which`, determine which indices in `var4` are assigned the `"MED"` level.

```{r error=T} 
which(mylist$var4 == "MED")

```

