---
title: "Homework #7"
author: "Name: Nicholas Ang"
output: html_document
---



## Question 1


Assume that `x` always takes a single integer between 0 and 9.
We wish to change `x` into a matching character string. For example, if the value of `x` is 3, then return `"three"`.
Use `switch()` function to produce a command that takes in `x` and returns a matching character string for all possible values $0, 1, \ldots, 9$.


```{r error=T} 
x <- 3
switch(x, "one","two","three","four","five","six","seven","eight","nine")
```


## Question 2


In lecture notes, we had 

```{r error=T} 
mystring <- "Bart"
x <- switch(mystring, Homer=12, Lisa=34, Bart=90, NA)
```

Now `mystring` is a character vector. Unfortunately, `switch()` function doesn't work for a vector.
For example, 

```{r error=T} 
mystring <- c("Bart", "Homer", "Mihye", "Homer")
x <- switch(mystring, Homer=12, Lisa=34, Bart=90, NA)
```

Thus, we wish to create a `for` loop to perform `switch()` function for *each element* of `mystring`. 

Note that `seq_along(mystring)` is the same as `1:length(string)`.
Since you need to perform `switch()` function for each element of `mystring`, you also need to save the output into a vector. 
To store your output, I initialized `myoutput` below. 
(Hint: You may use `myoutput[i]` to save your output from the `switch()` function at the i-th iteration)

Complete the following R code.


```{r error=T} 
mystring <- c("Bart", "Homer", "Mihye", "Homer")  # do not change this line
myoutput <- c()                                   # do not change this line

for(i in seq_along(mystring)){
  myoutput[i] = switch(mystring[i],  Homer=12, Lisa=34, Bart=90, NA)
}
print(myoutput)
```


## Question 3


__(a)__  Using a `for` loop, calculate $1+2+\cdots+x$ when $x$ is given.


```{r error=T} 
x <- 10
sumX = 0
for(i in 1:x){
  sumX = sumX + i
}
print(sumX)
```


__(b)__  Create a function to calculate $1+2+\cdots+x$ for any positive integer of `x`.


```{r error=T} 
Q3 <- function(x){
  # here, copy and paste your R code from part a (except the first line)
  sumX = 0
  for(i in 1:x){
    sumX = sumX + i
  }
  print(sumX)
}

# check if your function is correct
Q3(10)  # is it 55?
```


__(c)__  In part b, we didn't check whether `x` is a positive integer or not. 
Let's add an `if`-`else` statement to check whether `x` is positive or not. 
If `x` is not positive, return `"x must be positive"`.
(Note: We will not check whether `x` is an integer or not because it is not simple in R.)


```{r error=T} 
# copy and paste your Q3 function from part b, and modify it
Q3 <- function(x){
  # here, copy and paste your R code from part a (except the first line)
  if (x > 0)
  {
    sumX = 0
    for(i in 1:x){
      sumX = sumX + i
    }
    print(sumX)
  }else
  {
    print("x must be positive")
  }
}

# check if your function is correct
Q3(-5) 
```


## Question 4


Assume that the current stock price is 126.34.
Tomorrow's stock price will be changed based on the today's price. 
Thus, we multiplied a random number generated from Uniform distribution (0.99, 1.01).   
(This is just an assumption for HW, not reality!)

In the R code below, we used `repeat` statement to update a stock price everyday (by multiplying a random number). 

Now assume that if the stock price is below 124.5, it is a good time to buy more stocks.
Thus, while running `repeat` statement, we will wait until the stock price is below 124.5. 

* If the stock price is higher than or equal to 124.5, we will do nothing and wait the next day again. 

* If the stock price is finally below 124.5, we will return `"Stock price is below 124.5! Buy it while it's cheap!"` and stop running the `repeat` statement. 


```{r} 
set.seed(1026)

# Stock price
stock_price <- 126.34

repeat {
  # New stock price
  stock_price <- stock_price * runif(1, .99, 1.01)
  print(stock_price)
  
  # Check
  if(stock_price < 124.5) {
    print("Stock price is below 124.5! Buy it while it's cheap!")
    break
  }
}
```

Based on this R code, solve the following questions:


__(a)__  Change the condition in the `if` statement to break when the stock price is below 125. Do not forget to change the return message accordingly.


```{r error=T} 
set.seed(1026)

# Stock price
stock_price <- 126.34

repeat {
  # New stock price
  stock_price <- stock_price * runif(1, .99, 1.01)
  print(stock_price)
  
  # Check
  if(stock_price < 125) {
    print("Stock price is below 125! Buy it while it's cheap!")
    break
  }
}
```


__(b)__  Let's say that the current stock price is 67.55.  
In this question, we assume that a stock price decreases by 0.5\% everyday, thus multiply 0.995 to get the stock price for tomorrow. 
By using this rule, we will update the stock price everyday.

Fill in the blanks below to update the stock price everyday.
If the stock price is below 66, return `"Stock price is below 66! Buy it while it's cheap!"` and stop running the loop. 


```{r error=T} 
# Stock price
stock_price <- 67.55

repeat {
  # New stock price
  stock_price <- stock_price * .995
  print(stock_price)
  
  # Check
  if(stock_price < 66) {
    print("Stock price is below 66! Buy it while it's cheap!")
    break
  }
}
```


__(c)__  In part (b), how many times the stock price was printed? \
The stock price was printed 5 times.

__(d)__  Copy and paste your code from part (b). 
Now move the statement `print(stock_price)` to after the `if` statement, but still inside the loop. 


```{r error=T} 
# Stock price
stock_price <- 67.55

repeat {
  # New stock price
  stock_price <- stock_price * .995
  
  # Check
  if(stock_price < 66) {
    print("Stock price is below 66! Buy it while it's cheap!")
    break
  }
  print(stock_price)
}
```


__(e)__  In part (d), how many times the stock price was printed? \
The stock price was printed 4 times.


## Question 5


Consider a vehicle speed example. 


__(a)__  `speed` shows the current speed, and is initially set as 64. 
Complete the following `while` loop with the following characteristics:

* The condition of the `while` loop should check if `speed` is higher than 30.

* If `speed` is higher than 30, then print out `"Slow down!"`, decrease the speed *by* 7, and assign this new value to `speed` again. (if this step is wrong, you will get an infinite loop!)

* If `speed` is lower than or equal to 30, stop running the `while` loop. 


```{r error=T} 
speed <- 64

while (speed > 30) {
  print("Slow down!")
  speed <- speed - 7
}

print(speed)
```


__(b)__  In part (a), we checked whether `speed` is higher than 30 or not. Now we will update your code in part (a). 

Inside the body of the `while` loop, do the followings:

* If `speed` is higher than 48, then print out `"Slow down big time!"` and decrease `speed` by 11. 

* If `speed` is lower than or equal to 48, then print out `"Slow down!"` and decrease `speed` by 6. 


```{r error=T}  
speed <- 64

while (speed > 30) {
  print(paste("Your speed is",speed))
  if (speed > 48) {
    print("Slow down big time!")
    speed = speed - 11
  } else {
    print("Slow down!")
    speed = speed - 6
  }
}
```


__(c)__  In this question, we set the initial value of `speed` as 88. (it is too fast!)
Copy and paste your `if`-`else` statement from part (b). 

Let's add one more `if` statement now. 
Inside the `while` loop, if `speed` is higher than 80, then we don't want to continue running the `while` loop anymore. (The driver is out of control! We give up.) 
Fill in the blanks below.


```{r error=T} 
speed <- 88

while (speed > 30) {
  if (speed > 80) {
    print("The driver is out of control! We give up.")
    break
  }
  
  print(paste("Your speed is",speed))
  if (speed > 48) {
    print("Slow down big time!")
    speed = speed - 11
  } else {
    print("Slow down!")
    speed = speed - 6
  }
}
```


## Question 6


The initial value of `i` is 1. Complete the `while` loop so that it:

* prints out the triple of `i`, so `3*i` at each run.

* stops running if the triple of `i` is divisible by 8, but still prints out this triple before stopping.

To check whether `a` is divisible by `b` or not, you can use `a %% b == 0` in R.


```{r error=T} 
i <- 1

while (i <= 10) {
  print(3*i)
  if (3*i %% 8 == 0) {
    break
  }
  i <- i + 1
}
```


## Question 7


In HW6, we studied two ways to loop over a vector.

```{r} 
primes <- c(2, 3, 5, 7, 11, 13)

# example 1
for (p in primes) {
  print(p)
}

# example 2
for (i in 1:length(primes)) {
  print(primes[i])
}
```


__(a)__  Let's practice more. 
A vector `linkedin` contains the number of views your LinkedIn profile had in the last seven days.
Write a `for` loop that iterates over all the elements of `linkedin` and prints out every element separately.
Do this in two ways: 1) using the "example 1" and "example 2" above.


```{r error=T} 
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# example 1
for (i in linkedin){
  print(i)
}

# example 2
for (i in 1:length(linkedin)){
  print(linkedin[i])
}
```


__(b)__  Let's modify your first `for` loop in part (a). 
Add code to the `for` loop that loops over the elements of the `linkedin` vector:

* If the value of an element exceeds 10, print out `"You're popular!"`.

* If the value of an element does not exceed 10, print out `"Be more visible!"`.


```{r error=T} 
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

for (i in linkedin) {
  if (i > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  print(i)
}
```


__(c)__  Let's extend your `for` loop more. Copy and paste your `if`-`else` statement from part (b). 
After that, add two more `if` statements as follows:

* If the value of an element exceeds 16, print out `"This is ridiculous. I am outta here!"` and stop running the `for` loop.

* If the value of an element is lower than 5, print out `"This is too embrrasing!"` and move to the next iteration. 


```{r error=T} 
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

for (i in linkedin) {
  if (i > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  
  # Add if statement
  if(i > 16){
    print("This is ridiculous. I am outta here!")
    break
  }
  
  # Add if statement
  if(i < 5){
    print("This is too embarrassing!")
    next
  }
  
  print(i)
}
```


## Question 8


We have a matrix `A` defined as below.


```{r} 
A <- matrix(1:12, 3, 4)
print(A)
```


Print out each element of the matrix `A`. That is, print out (1,1)th element, (1,2)th element, (1,3)th element, (1,4)th element, (2,1)th element, and so on, in this order.

To do this, you need a `for` loop inside a `for` loop. 

* The outer loop should loop over the rows, with loop index `i`.

* The inner loop should loop over the columns, with loop index `j`.

* Inside the inner loop, use `print()` function to print out the (i,j)th element of the matrix `A`.


```{r error=T} 
for (i in 1:nrow(A)) {
  for (j in 1:ncol(A)) {
    print(A[i,j])
  }
}
```


## Question 9


In lecture notes, we defined a function `myfactorial` to calculate $x!$ as below:

```{r error=T} 
myfactorial <- function(x){
  i <- 1
  temp <- 1
  while(i<=x){
    temp <- temp*i
    i <- i+1
  }
  return(temp)
}
```

In fact, this is not a good function. It works okay when the value of `x` is non-negative. 
However, when `x` is negative, it still gives us the value of `1` which is not true.
For example,

```{r error=T} 
myfactorial(-2)
```

Let's modify the function below as the following:

* Check whether the value of `x` is non-negative or not. 

* If it is non-negative, go ahead to perform the body code in the braced area (from the 2nd to 8th line).

* If it is negative, print out `"x is a negative value."` and do not perform the body code in the braced area.  


```{r error=T} 
myfactorial <- function(x){
  if(x > 0)
  {
    i <- 1
    temp <- 1
    while(i<=x){
      temp <- temp*i
      i <- i+1
    }
    return(temp)
  }else
  {
    print("x is a negative value")
  }
}


# check if it prints out "x is a negative value."
myfactorial(-2)  
```


## Question 10


Create a function named `pwr` as follows:

* It raises a number to any power (that is, $x^y$). 

* Make the default raise the input number(s) to the power of 2.

* To check if it works fine, calculate $3^5$ using your function. 


```{r error=T} 
pwr = function(x,y = 2){
  x^y
}

pwr(3,5)

```


## Question 11  


The probability density function of a log-normal distribution with $\mu=1$ and $\sigma^2=1$ is given by

\[
f(x)=\frac{1}{x\sqrt{2\pi}}\exp(-\frac{(\log(x)-1)^2}{2})
\]


__(a)__  Create a function named `lognormal` to calculate the value of $f(x)$ for any $x$.


```{r error=T} 
lognormal = function(x){
  fx = (1/(x*sqrt(2*pi))  * exp(-((log(x, base = exp(1)) - 1)^2)/2))
  return(fx)
}
```


__(b)__  Using `lognormal` function defined in part (a), calculate the value of $f(5)$. Is it 0.06626564?


```{r error=T} 
f5 = lognormal(5)
print(f5)
```
Yes, it is 0.06626564


## Question 12


In HW3, we had three vectors of Apple, IBM, and Microsoft stock prices and combined them as `stocks`.


```{r} 
apple <- c(109.49, 109.90, 109.11, 109.95, 111.03, 112.12, 113.95, 113.30, 115.19, 115.19,115.82, 115.97, 116.64, 116.95, 117.06, 116.29, 116.52, 117.26, 116.76, 116.73, 115.82)

ibm <- c(159.82, 160.02, 159.84, 160.35, 164.79, 165.36, 166.52, 165.50, 168.29, 168.51, 168.02, 166.73, 166.68, 167.60, 167.33, 167.06, 166.71, 167.14, 166.19, 166.60, 165.99)

micr <- c(59.20, 59.25, 60.22, 59.95, 61.37, 61.01, 61.97, 62.17, 62.98, 62.68, 62.58, 62.30, 63.62, 63.54, 63.54, 63.55, 63.24, 63.28, 62.99, 62.90, 62.14)

stocks <- cbind(apple, ibm, micr)
head(stocks)

# Do not use vectors apple, ibm, and micr.
rm(list=c("apple", "ibm", "micr"))
```


__(a)__  Let's calculate the mean and sum for each variable in `stocks` and store them in a matrix `B`. 
`B` is a $2\times 3$ matrix. 
In the first row, we store the mean for each variable. 
In the second row, we store the sum for each variable. 
The first column is for `apple`, the second column is for `ibm`, and the third column is for `micr`. 
That is, 

\[
B = \left(\begin{array}{ccc}
\mbox{mean(apple)} & \mbox{mean(ibm)} & \mbox{mean(micr)} \\ \mbox{var(apple)} & \mbox{var(ibm)} & \mbox{var(micr)}
\end{array} \right)
\]

Complete the following `for` loop. 

Hint: Calculate the mean and sum for the $i$th column vector of `stocks` and save those two values into the $(1,i)$th and $(2,i)$th elements of `B`, respectively.
We have three columns in `stocks`. Thus, `i` is from 1 to 3.


```{r error=T} 
# initialize
B <- matrix(NA, 2, 3) 

for (i in 1:ncol(stocks)){
    B[1, i] <- mean(stocks[, i])
    B[2, i] <- sum(stocks[, i])
}
print(B)
```


__(b)__  In part (a), we calculated the mean and sum for each column separately. 
Now let's calculate the column mean and column sum.  By doing so, you don't need to repeat the calculation for each column.

* Calculate the column mean of `stocks` and save them into the first row of `B`,

* Calculate the column sum of `stocks` and save them into the second row of `B`,


```{r error=T} 
# initialize
B <- matrix(NA, 2, 3) 

B[1,] <- colMeans(stocks)    # column mean
B[2,] <- colSums(stocks)    # column sum

print(B)
```

Have you got the same output as in part (a)?


__(c)__  Now we have a missing value in `stocks`. 

```{r} 
stocks[12,1] <- NA
```

Copy and paste your code from part (a). It doesn't work anymore due to a missing value.
To calculate the mean and sum, modify your code below.


```{r error=T}  
# initialize
B <- matrix(NA, 2, 3) 

for (i in 1:ncol(stocks)){
    B[1, i] <- mean(stocks[, i], na.rm = T)
    B[2, i] <- sum(stocks[, i], na.rm = T)
}
print(B)
```


__(d)__  You still have a missing value in `stocks`. 
Copy and paste your code from part (b). It doesn't work anymore due to a missing value.
To calculate the column mean and column sum, modify your code below.


```{r error=T}  
# initialize
B <- matrix(NA, 2, 3) 

B[1,] <- colMeans(stocks, na.rm = T)    # column mean
B[2,] <- colSums(stocks, na.rm = T)    # column sum

print(B)
```


## Question 13


When we have a sample data $X_1,X_2,\ldots,X_n$, a sample variance is defined by
\[
\frac{1}{n-1}\sum_{i=1}^n (X_i-\overline{X})^2,
\]
where $\overline{X}$ is a sample mean and $n$ is a sample size. 


__(a)__  There is a built-in function `var` in R.
Create your own function `variance` to calculate a sample variance. 
To make it simple, you are allowed to use other built-in functions, such as `mean`, `sum`, `length`. 
Assume that a sample data is stored in `x` as a vector.


```{r error=T}  
variance <- function(x){
  sumX = 0
  meanX = mean(x)
  (1/(length(x)-1)) * sum((x - meanX)^2)
}
```


__(b)__  Let's check if your function works. A sample data `x` is given below.
Calculate the sample variance of `x` using `variance` function defined in part (a).


```{r error=T}  
x <- c(5, 2, 6, 2, 7, 4)
variance(x)
```


## Question 14


Let $h(x,n) = 1+x+x^2+\cdots+x^n = \sum_{i=0}^n x^i$. 


__(a)__  Create a function `h` to calculate $h(x,n)$ using a `for` loop.


```{r error=T}  
h <- function(x, n){
  xi = 1
  for(i in 1:n)
  {
    xi = xi + x^i
  }
  return(xi)
}
```


__(b)__  Using `h` function, calculate $h(2,5)$. Is it 63?


```{r error=T}  
h(2,5)
```
Yes, it is 63


## Question 15


When we have a sample data $X_1,X_2,\ldots,X_n$, a geometric mean is defined by
\[
\Big(\prod_{i=1}^n X_i\Big)^{1/n}
\]


__(a)__  Create a function `geometric` to calculate a geometric mean.
Assume that a sample data is stored in `x` as a vector. Fill in the blanks below.


```{r error=T}  
geometric <- function(x){
  temp <- 1             # initialize
  
  for(i in 1:length(x)){    # loop over all elements of x
    temp <- temp * x[i]
  }
  
  final <- (temp)^(1/length(x))
  return(final)
}
```


__(b)__  Let's check if your function works. A sample data `x` is given below.
Calculate the geometric mean of `x` using `geometric` function defined in part (a).
Is it 3.870109?


```{r error=T}  
x <- c(5, 2, 6, 2, 7, 4)
geometric(x)
```
Yes, it is 3.870109


## Question 16


The standard form of a quadratic equation is
\[
ax^2 + bx + c =0,
\]
where $a$, $b$, and $c$ are known values and $a\neq 0$.

The quadratic formula is:
\[
x = \frac{-b\pm \sqrt{b^2-4ac}}{2a}
\]

When the discriminant $b^2-4ac$ is

* positive, there are two real solutions.

* zero, there is one real solution.

* negative, there are two complex solutions.


__(a)__  Let's create a function `quadratic` to solve a quadratic equation as follows:

* If `a` is zero, print out `"This is not a quadratic equation"` and do not calculate anything.

* If the discriminant $b^2-4ac$ is positive, return two solutions. 

* If the discriminant $b^2-4ac$ is zero, return one solution. The solution is simply $x=\frac{-b}{2a}$.

* If the discriminant $b^2-4ac$ is negative, do not calculation anything else and simply print out `"It's complex!"`. 


```{r error=T}  
quadratic <- function(a,b,c){
  if(a == 0)
  {
    print("This is not a quadratic equation")
  }else{
    if(b^2 - 4*a*c > 0)
    {
      x1 = (-b + sqrt(b^2 - (4*a*c)))/(2*a)
      x2 = (-b - sqrt(b^2 - (4*a*c)))/(2*a)
      return(c(x1,x2))
    }else if(b^2 - 4*a*c == 0){
      x = -b/(2*a)
      return(x)
    }else
    {
      print("It's complex!")
    }
  }
}
```


__(b)__  Using `quadratic` function defined in part (a), solve $2x^2 + 3x + 1 = 0$


```{r error=T}  
quadratic(2,3,1)
```



## Question 17


Suppose that $X_1, \ldots, X_n$ is a sample from a normal distribution having unknown parameters $\mu$ and $\sigma^2$.
Let's perform a hypothesis test for a population variance $\sigma^2$.  
Unfortunately, there is no built-in function for this test.

In this example,

* Null hypothesis: $H_0: \sigma^2 = \sigma_0^2$

* Alternative hypothesis: $H_1: \sigma^2 > \sigma_0^2$

* Test statistic: $\displaystyle\frac{(n-1)S^2}{\sigma_0^2}$

* If a p-value is less than $\alpha$, we reject $H_0$.


__(a)__  Create a function `mytest` to perform a hypothesis test for a population variance. 
Similar to z-test in lecture notes, fill in the blanks below.

In this function, x is a data vector, `sigma02` is $\sigma_0^2$, $S^2$ can be calculated by `var(x)`, and $\alpha$ is a significance level. 

(I calculated a p-value for you. You don't need to modify it.)


```{r error=T}  
mytest <- function(x, sigma02, alpha){        # input: x, sigma02, alpha
  
  H0 <- paste("HO: sigma^2 =", sigma02)           # null hypothesis
  H1 <- paste("H1: sigma^2 >", sigma02)           # alternative hypothesis
  
  n <- length(x)                              # n = sample size
  chi2 <- ((n-1)*var(x))/sigma02                # test statistic
  
  p <- pchisq(chi2, n-1, lower.tail=F)        # p-value
  
  if (p < alpha) {                              # p-value is less than alpha?
    result <- "Reject the null hypothesis"
  } else{
    result <- "Fail to reject the null hypothesis"
  }
  
  # return
  
  list(hypotheses = c(H0,H1),                 # hypotheses
       test.statistic = chi2,              # test statistic
       p.value = p,                     # p-value
       test.result = result)                # show result of hypothesis test
}
```


__(b)__  Using `mytest` function defined in part (a), perform a hypothesis test for a population variance.
In this example, the data `x` is given below.
The null and alternative hypotheses are $H_0: \sigma^2 = 2$ and $H_1: \sigma^2 > 2$. 
Use $\alpha=0.05$. 


```{r error=T}  
x <- c(61.1,58.2,62.3,64.0,59.7,66.2,57.8,61.4)
mytest(x,2,0.05)
```


## Question 18


Suppose you are tasked with computing the precise dosage amounts of a certain drug in a collection of hypothetical scientific experiments. 

These amounts depend upon some predetermined set of "dosage thresholds" (`lowdose`, `meddose`, and `highdose`), as well as a predetermined dose level factor vector named `doselevel`. 

Write a set of nested `if` statements that produce a new numeric vector called `dosage`, according to the following rules:

* First, *if* there are any instances of `"High"` in `doselevel`, perform the following operations:

  + Check *if* `lowdose` is greater than or equal to 10. If so, overwrite `lowdose` with 10; 
    *otherwise*, overwrite `lowdose` by itself divided by 2.

  + Check *if* `meddose` is greater than or equal to 26. If so, overwrite `meddose` by 26.
  
  + Check *if* `highdose` is less than 60. If so, overwrite `highdose` with 60; 
    *otherwise*, overwrite `highdose` by itself multiplied by 1.5.

  + Create a vector named `dosage` with the value of `lowdose` repeated (`rep`) to match the `length` of `doselevel`.

  + Overwrite the elements in `dosage` corresponding to the index positions of instances of `"Med"` in `doselevel` by `meddose`.
  
  + Overwrite the elements in `dosage` corresponding to the index positions of instances of `"High"` in `doselevel` by `highdose`.

* *Otherwise* (in other words, if there are no instances of `"High"` in `doselevel`), perform the following operations:

  + Create a new version of `doselevel`, a factor vector with levels `"Low"` and `"Med"` only, and label these with `"Small"` and `"Large"`, respectively .

  + Check to see *if* `lowdose` is less than 15 AND `meddose` is less than 35. If so, overwrite `lowdose` by itself multiplied by 2 and overwrite `meddose` by itself plus `highdose`.
  
  + Create a vector named `dosage`, which is the value of `lowdose` repeated (`rep`) to match the `length` of `doselevel`.

  + Overwrite the elements in `dosage` corresponding to the index positions of instances of `"Large"` in `doselevel` by `meddose`.


```{r error=T} 
# example
lowdose <- 12.5
meddose <- 25.3
highdose <- 58.1
doselevel <- factor(c("Low","High","High","High","Low","Med", "Med"),levels=c("Low","Med","High"))

# there are any instances of "High" in doselevel?
if (any(doselevel == "High")){
  
  # lowdose is greater than or equal to 10 ?
  if (lowdose >= 10){
    
    # overwrite lowdose with 10
    lowdose <- 10
      
  } else{
    # overwrite lowdose by itself divided by 2
    lowdose <- lowdose/2
  }
  
  # meddose is greater than or equal to 26 ?
  if (meddose >= 26){
    
    # overwrite meddose by 26
    meddose <- 26
  } 
  
  # highdose is less than 60 ?
  if (highdose < 60){
    
    # overwrite highdose with 60
    highdose <- 60
      
  } else{
    # overwrite highdose by itself multiplied by 1.5
    highdose <- highdose*1.5
  }
  
  # Create dosage with the value of lowdose repeated to match the length of doselevel
  dosage <- rep(lowdose, length(doselevel))

  # Overwrite the elements in dosage corresponding to the index positions of instances of "Med" in doselevel by meddose
  dosage[doselevel == "Med"] <- meddose
  
  # Overwrite the elements in dosage corresponding to the index positions of instances of "High" in doselevel by highdose
  dosage[doselevel == "High"] <- highdose
    
} else{
  # Create a new version of doselevel with levels "Low" and "Med" only, and label these with "Small" and "Large", respectively
  doselevel <- factor(doselevel, c("Low", "Med"), c("Small", "Large"))

  # lowdose is less than 15 AND meddose is less than 35 ?. If so, 
  if (lowdose < 15 & meddose < 35){
    
    # overwrite lowdose by itself multiplied by 2
    lowdose <- lowdose*2
    
    # overwrite meddose by itself plus highdose
    meddose <- meddose + highdose
  }
  
  # Create dosage, which is the value of lowdose repeated to match the length of doselevel
  dosage <- rep(lowdose, length(doselevel))

  # Overwrite the elements in dosage corresponding to the index positions of instances of "Large" in doselevel by meddose
  dosage[doselevel == "Large"] <- meddose
  
}
print(dosage)
```



